---
title: "Accessbility to possible sites"
output: html_notebook
author: Mark Green
---

__Aim:__ To calculate accessbility to all possible sites within Liverpool.

In the notebook, I have calculated the 15 minute walking distance for all possible venues to give an idea of where to select the best locations. I have done this for all the sites Captain Richard Fergus sent to me today. 7 no recce conducted sites were removed as they contained no latitude or longitude records. The colours of the site coverage follows the same as in the excel file, so:

* Active = darker green (n=71)
* Approved = lighter green (n=1)
* Follow up required  = yellow (n=7)
* No recce conducted = white/grey (n=30)
* Suitable = blue (n=15)
* Unsuitable = red (n=91)

You can turn off and on the each of these layers by clicking on them in the right hand box.

```{r message=FALSE, warning=FALSE}
# Libraries
library(maxcovr) # For Location-Allocation modelling
library(dplyr) # For data wrangling
library(purrr)
library(leaflet) # For mapping
library(sf)
library(ggplot2) # For plotting
library(hereR) # For routing

# Set API key for 'here'
set_key("IXVUkHClCXqgBf3lJeGyJij0TfU6zG5mInbC-uaCtn0") # You will need to register and request a key - is free

# Load data
sites <- read.csv("./all_possible_sites_small.csv") # Location of all active and inactive test sites 

# Convert sites to spatial points format
sites_sp <- sites %>% 
  st_as_sf(coords = c("long", "lat")) %>% # Define co-ordinates (or X / Y)
  st_set_crs(4326)

# Subset by type
active <- sites_sp[sites_sp$state == "Active",]
approved <- sites_sp[sites_sp$state == "Approved",]
follow_up <- sites_sp[sites_sp$state == "Follow up required",]
no_recce <- sites_sp[sites_sp$state == "No Recce Conducted",]
suitable <- sites_sp[sites_sp$state == "Suitable - Awaiting Approval",]
unsuitable <- sites_sp[sites_sp$state == "Unsuitable",]

# Calculate walking accessibility around each test site state
active_access <- isoline(active, mode = "pedestrian", range = 15*60, range_type = "time") # 15 min walk distance
approved_access <- isoline(approved, mode = "pedestrian", range = 15*60, range_type = "time") # 15 min walk distance
follow_up_access <- isoline(follow_up, mode = "pedestrian", range = 15*60, range_type = "time") # 15 min walk distance
no_recce_access <- isoline(no_recce, mode = "pedestrian", range = 15*60, range_type = "time") # 15 min walk distance
suitable_access <- isoline(suitable, mode = "pedestrian", range = 15*60, range_type = "time") # 15 min walk distance
unsuitable_access <- isoline(unsuitable, mode = "pedestrian", range = 15*60, range_type = "time") # 15 min walk distance

# Convert to polygon format for mapping
active_access_p <- active_access %>%
                        st_cast("POLYGON")
approved_access_p <- approved_access %>%
                        st_cast("POLYGON")
follow_up_access_p <- follow_up_access %>%
                        st_cast("POLYGON")
no_recce_access_p <- no_recce_access %>%
                        st_cast("POLYGON")
suitable_access_p <- suitable_access %>%
                        st_cast("POLYGON")
unsuitable_access_p <- unsuitable_access %>%
                        st_cast("POLYGON")

# Map existing sites
leaflet() %>%
  addPolygons(data = active_access_p, 
              group = "Active",
              color = "green") %>% # Active sites
    addPolygons(data = approved_access_p,
              group = "Approved",
              color = "palegreen") %>% # Approved sites
    addPolygons(data = follow_up_access_p,
              group = "Follow up",
              color = "yellow") %>% # Follow up sites
    addPolygons(data = no_recce_access_p,
              group = "No recce",
              color = "snow2") %>% # No recce sites
    addPolygons(data = suitable_access_p,
              group = "Suitable",
              color = "steelblue") %>% # Suitable sites
    addPolygons(data = unsuitable_access_p,
              group = "Unsuitable",
              color = "orangered") %>% # Unsuitable sites
  addProviderTiles("CartoDB.Positron", 
                   group = "Base") %>%
  addLayersControl(#baseGroups = c("Base"), # Control layers in map
    overlayGroups = c("Active", "Approved", "Follow up", "No recce", "Suitable", "Unsuitable"),
    options = layersControlOptions(collapsed = FALSE))




```

I have also reproduced this map for the active sites that are accessible to the public and advertised on the website (e.g. the University of Liverpool or City of Liverpool College sites are only accessbile to those who attend those sites). Again we map the same, but only consider the layers that might be possible in the time available.

```{r}
# Load data
sites <- read.csv("./ltc.csv") # Location of test sites (data taken from the whole city testing website, happy to share)

# Convert sites to spatial points format
sites_sp <- sites %>% 
  st_as_sf(coords = c("X", "Y")) %>% # Define co-ordinates
  st_set_crs(4326)

# Calculate walking accessibility around each test site
access <- isoline(sites_sp, mode = "pedestrian", range = 15*60, range_type = "time") # 15 min walk distance

# Convert to polygon format for mapping
access_p <- access %>%
     st_cast("POLYGON")

# Map
leaflet() %>%
     addPolygons(data = suitable_access_p,
                 group = "Suitable",
                 color = "steelblue") %>% # Suitable sites
      addPolygons(data = approved_access_p,
              group = "Approved",
              color = "palegreen") %>% # Approved sites
    addPolygons(data = follow_up_access_p,
              group = "Follow up",
              color = "yellow") %>% # Follow up sites
    addPolygons(data = no_recce_access_p,
              group = "No recce",
              color = "snow2") %>% # No recce sites
     addPolygons(data = access_p,
                 group = "Active to public",
                 color = "green") %>% # Plot 15 min walkability
     addProviderTiles("CartoDB.Positron",
                      group = "Base") %>%
    addLayersControl(#baseGroups = c("Base"), # Control layers in map
    overlayGroups = c("Active to public", "Approved", "Follow up", "No recce", "Suitable"),
    options = layersControlOptions(collapsed = FALSE))


```

The takeaway for me is that some of the no recce sites appear to be in good locations for improving geographic coverage (assuming we have the time to investigate them). The suitable sites are in decent locations, but might cover some of the same ground.

We can identify the 12 best sites that give the best geographical coverage (whether you wish to target certain areas as well, I cannot say from here). We will just consider those sites that are approved, need follow up, suitable or no recce in the analysis. We compare accessibility to the above map (i.e. only those accessible to the public).

```{r}
# Load data
oas <- read_sf(dsn = "./liverpool_oa/england_oa_2011.shp") # Output areas (data from UKBorders, happy to share)
# st_crs(oas) # Check CRS - see 'user input' at top
oas <- st_transform(oas, 4326) # Transform to same CRS as lat/long

# Calculate centroid of output areas
oa_centres <- st_centroid(oas) # Find centroid
oas_sf <- st_as_sf(oa_centres, coords = c("lon","lat")) %>% # Re-format data
  st_set_crs(4326)
oas_coords <- unlist(st_geometry(oas_sf)) %>% # Convert to Lat/Long
  matrix(ncol=2,byrow=TRUE) %>% 
  as_tibble() %>% 
  setNames(c("lon","lat"))

# # Describe current coverage
# coverage(current_sites, 
#          oas_coords,
#          distance_cutoff = 100)

# Edit inputs to match required format
current_sites <- sites
current_sites$Location <- NULL
names(current_sites)[names(current_sites)=="X"] <- "long"
names(current_sites)[names(current_sites)=="Y"] <- "lat"
names(oas_coords)[names(oas_coords)=="lon"] <- "long"

# Load in possible new test site locations
poss_sites <- sites[sites$state == "Approved" | 
                      sites$state == "Follow up required" | 
                      sites$state == "Suitable - Awaiting Approval" | 
                      sites$state == "No Recce Conducted",]
pcs <- poss_sites[,c("lat", "long")] # Subset all postcodes to match required format

# Run location-allocation model to select five best new sites
mc_12 <- max_coverage(existing_facility = current_sites, # List of current test sites
                      proposed_facility = pcs, # List of all possible postcodes
                      user = oas_coords, # Data to optimise to (i.e. output area centroids)
                      n_added = 12, # How many facilities
                      distance_cutoff = 800) # 800m bufer radius 
#summary(mc_12) # Plot summary statistics
#mc_12$facility_selected[[1]] # List sites selected
temp <- merge(data.frame(mc_12$facility_selected[[1]]), sites, by.x = c("lat", "long"), by.y = c("lat", "long"), all.x = TRUE) # Join onto names of sites
temp

```

They are the best 12 sites (there are 14 sites here, but only because some lat/longs share multiple sites). We can also map them too.

```{r}
# Map new and existing sites
leaflet() %>%
  ## Plot exsting site locations ##
  addPolygons(data = access_p) %>% # Plot 15 min walkability
  ## Plot proposed new site locations ##
  addCircleMarkers(data = mc_12$facility_selected[[1]], # Data for plotting new sites as points
                   radius = 1, 
                   color = "green") %>% 
  addCircles(data = mc_12$facility_selected[[1]], 
             radius = 800,
             fill = FALSE,
             weight = 3,
             color = "green") %>%
  addProviderTiles("CartoDB.Positron") # Add on map layer below

```

